|

### 2021-12-16 05-Chp-CPRD-Analysis.Rmd
End of day:

Little evidence for an effect of lipid-regulating agents on the risk of Alzheimer's disease was found, with the exception of a slightly increased risk in those prescribed a certain type of lipid-regulating agent called fibrates. In contrast, I found an increased risk of vascular and other (i.e. non-Alzheimer's) dementia was associated with lipid-regulating agent use.

@@

This increased risk in outcomes with a vascular element (e.g. vascular dementia) is unexpected, and is very likely to be due to the presence of bias in my analysis. This bias, called "confounding by indication", is caused when those who are prescribed a statin are more at risk of vascular dementia for a range of reasons, which makes it appear as if statins are harmful. Despite this limitation, the analysis provides an important source of information which will be used in later chapters.

@@

Previously known as the General Practice Research Database, the Clinical Practice Research Datalink (CPRD) is a large population-based EHR database.[@herrett2015] The database has been collecting primary care data from participating practices across England since 1987.[@williams2012; @wood2001revitalizing] It contains the records of more than 10 million primary care patients in England, and is broadly representative of the UK population in terms of age, sex and ethnicity.[@herrett2015; @mathur2014]

To avoid the ambiguity of interpreting free-text clinical notes and to allow for easy analysis of the resulting data, the CPRD primarily collects data using a predefined coding system known as Read codes.[@booth1994] All clinical events, including  test results and diagnoses, can be identified by a specific Read code. The codes use a nested approach (see Table \@ref(tab:readExample-table)), with the initial characters defining broad diagnostic topics (e.g. Eu... - Mental and behavioural disorders), while subsequent characters provide additional information on the specific condition diagnosed (e.g. Eu001 - Dementia in Alzheimer's disease with late onset).

@@

* a Read code for a diagnosis of hypercholesterolemia or related condition; 
* a Read code for prescription of a lipid-regulating agent (such as statins); 
* a total cholesterol test result of >4 mmol/L; 
* an LDL-c test result of >2 mmol/L.

@@

* an outcome of interest;
* death; 
* end of follow-up (29 February 2016);
* last registration date with their GP practice;
* the last CPRD collection date for their practice.

Participants were ineligible for the cohort if they were less than 40 years of age (because these patients are less likely to be prescribed a LRA), had less than 12 months of "research quality" data, were simultaneously prescribed more than one lipid-regulating agent (due to the difficultly of assigning these to a single exposure group), or were diagnosed with an outcome of interest before or on the date of the index event (i.e. had less than one full day of follow-up).

@@

A participants drug class was assigned based on their first recorded prescription, and any drug switching was ignored in an effort to mimic an intention-to-treat approach. I did however examine how often the initial drug class altered according to one of three criteria:

@@

Missing data was only considered problematic for variables where a numerical test result was expected (e.g. BMI), or where a code existed for the absence of the condition (e.g. categorical smoking status). This approach was necessary because absence of a code for other treatments or conditions (e.g. statin use or dementia) was assumed to indicate absence of the treatment/condition rather than being considered missing.[@wells2013strategies]

@@

* $t$ is the survival time;
* $h(t)$ is the hazard function;
* $x_1,x_2,...,x_p$ are the covariates which determine the hazard function, while $b_1,b_2,...,b_p$ are the coefficients for each covariate.; and
* $h_o(t)$ is the baseline hazard - when all $x_i$ are zero, the $exp()$ function resolves to 1.

@@

Immortal time bias describes two distinct but related types of bias, considered here in relation to statin use (Figure \@ref(fig:immortalTimeBias)). The first presentation, the selection bias aspect (Panel A), occurs when time prior to statin initiation is excluded leading to the statin and control groups being followed up from different time points.[@levesque2010] The unexposed group are followed from the date of an index event (e.g. diagnosis of hypercholesterolemia), while the statin group is followed from date of a statin initiation. In this scenario, the time between the index event and statin initiation is missing, and any events that occur in the exposed group prior to the prescription will be inappropriately excluded from the analysis.

@@

Patients start in the unexposed group and contribute time-at-risk until they are prescribed a lipid-regulating agent, at which point they move into the exposed group. Note, patients for whom prescription of a lipid-regulating agent was the index event only contribute time to the exposed group (i.e. they enter the cohort and move into the exposed group on the same day).

@@

To observe the effect of adjusting for additional covariates, I ran two additional models unadjusted except for: (a) age; and (b) age and gender. The results of these models was then compared the results with the fully adjusted model.

@@

My results show litte evidence was found for an effect of lipid-regulating agents on probable (`r probad_text`) and possible (`r possad_text`) Alzheimer's disease when compared to no treatment, with the sole exception of fibrates on probable Alzheimer's disease (`r probad_fib_text`).

@@

In almost all cases, the use of imputed data resulted in a marginal attenuation of the effects observed when using a complete case analysis. It should be noted that due to the large amount of missing data (e.g. `r ldltext` were missing a baseline LDL cholesterol measure), the number of participants included in the complete case analysis was substantially smaller than that included when using imputed data. In this case, though the overall position of the effect estimates does not change substantially when using the imputed dataset, there is a noticeable gain in power.[@sterne2009]

@@

Lipid-regulating agents showed little evidence of an association with probable and possible Alzheimer's disease when compared to no treatment, but were associated with increased risk of an all-cause dementia, vascular dementia and other dementias diagnoses. The estimate observed in each case was driven by the effects observed in the statin subgroup because a substantial majority of participants prescribed an LRA were prescribed a statin. For the other drug classes, no association was found with any outcome, with two exceptions. Ezetimibe was associated with increased risk of vascular and other dementias, while fibrates were associated with increase risk of all-cause dementia and probable Alzheimer's disease.

@@

<!----------------------------------------------------------------------------->
(ref:smeethComparison-cap) __Comparison of codelists__ - Sankey diagram comparing the codes used in this analysis with those used in the Smeeth _et al_ paper.[@smeeth2009] The outcomes and number of codes contributing to each are presented (the Smeeth _et al_ outcomes are on the right hand side of the figure). The joining lines showing the overlap between the categories in the two analyses.

(ref:smeethComparison-scap) Comparison of codelists

@@

This analysis serves to illustrative the importance of the code lists chosen to define the outcomes of interest in EHR, particularly if they are used to define competing outcomes (e.g. AD vs non-AD dementia). The different codes used by Smeeth _et al._, in addition to an analytical approach that adjusted for covariates defined after the index date, may go some way to explaining why my analysis obtained different results despite the substantial overlap in the data sources used.

@@

* I found little evidence for an effect of lipid-regulating agents on probable or possible Alzheimer’s disease. However, lipid-regulating agent use was associated with an increased risk of all-cause, vascular and other dementias. In all cases, the estimated associations for the "any LRA" analyses were driven by those those observed in the large statin subgroup.

>>>>

Little evidence for an effect of lipid-regulating agents on the risk of Alzheimer's disease was found, with the exception of a slightly increased risk in those prescribed a certain type of lipid-regulating agent called fibrates. In contrast, I found an increased risk of vascular and other (i.e., non-Alzheimer's) dementia was associated with lipid-regulating agent use.

@@

This increased risk in outcomes with a vascular element (e.g., vascular dementia) is unexpected and is very likely to be due to the presence of bias in my analysis. This bias, called "confounding by indication", is caused when those who are prescribed a statin are more at risk of vascular dementia for a range of reasons, which makes it appear as if statins are harmful. Despite this limitation, the analysis provides an important source of information which will be used in later chapters.

@@

Previously known as the General Practice Research Database, the Clinical Practice Research Datalink (CPRD) is a large population-based EHR database.[@herrett2015] The database has been collecting primary care data from participating practices across England since 1987.[@williams2012; @wood2001revitalizing] It contains the records of more than 10 million primary care patients in England and is broadly representative of the UK population in terms of age, sex and ethnicity.[@herrett2015; @mathur2014]

To avoid the ambiguity of interpreting free-text clinical notes and to allow for easy analysis of the resulting data, the CPRD primarily collects data using a predefined coding system known as Read codes.[@booth1994] All clinical events, including  test results and diagnoses, can be identified by a specific Read code. The codes use a nested approach (see Table \@ref(tab:readExample-table)), with the initial characters defining broad diagnostic topics (e.g., Eu... - Mental and behavioural disorders), while subsequent characters provide additional information on the specific condition diagnosed (e.g., Eu001 - Dementia in Alzheimer's disease with late onset).

@@

* a Read code for a diagnosis of hypercholesterolemia or related condition; 
* a Read code for prescription of a lipid-regulating agent (such as statins); 
* a total cholesterol test result of >4 mmol/L; or
* an LDL-c test result of >2 mmol/L.

@@

* an outcome of interest;
* death; 
* end of follow-up (29 February 2016);
* last registration date with their GP practice; or
* the last CPRD collection date for their practice.

Participants were ineligible for the cohort if they were less than 40 years of age (because these patients are less likely to be prescribed a LRA), had less than 12 months of "research quality" data, were simultaneously prescribed more than one lipid-regulating agent (due to the difficultly of assigning these to a single exposure group), or were diagnosed with an outcome of interest before or on the date of the index event (i.e., had less than one full day of follow-up).

@@

A participant's drug class was assigned based on their first recorded prescription, and any drug switching was ignored in an effort to mimic an intention-to-treat approach. I did however examine how often the initial drug class altered according to one of three criteria:

@@

Missing data was only considered problematic for variables where a numerical test result was expected (e.g., BMI), or where a code existed for the absence of the condition (e.g., categorical smoking status). This approach was necessary because absence of a code for other treatments or conditions (e.g., statin use or dementia) was assumed to indicate absence of the treatment/condition rather than being considered missing.[@wells2013strategies]

@@

* $t$ is the survival time;
* $h(t)$ is the hazard function;
* $x_1,x_2,...,x_p$ are the covariates which determine the hazard function, while $b_1,b_2,...,b_p$ are the coefficients for each covariate; and
* $h_o(t)$ is the baseline hazard - when all $x_i$ are zero, the $exp()$ function resolves to 1.

@@

Immortal time bias describes two distinct but related types of bias, considered here in relation to statin use (Figure \@ref(fig:immortalTimeBias)). The first presentation, the selection bias aspect (Panel A), occurs when time prior to statin initiation is excluded leading to the statin and control groups being followed up from different time points.[@levesque2010] The unexposed group are followed from the date of an index event (e.g., diagnosis of hypercholesterolemia), while the statin group is followed from date of a statin initiation. In this scenario, the time between the index event and statin initiation is missing, and any events that occur in the exposed group prior to the prescription will be inappropriately excluded from the analysis.

@@

Patients start in the unexposed group and contribute time-at-risk until they are prescribed a lipid-regulating agent, at which point they move into the exposed group. Note, patients for whom prescription of a lipid-regulating agent was the index event only contribute time to the exposed group (i.e., they enter the cohort and move into the exposed group on the same day).

@@

To observe the effect of adjusting for additional covariates, I ran two additional models unadjusted except for: (a) age; and (b) age and gender. The results of these models were then compared the results with the fully adjusted model.

@@

My results show little evidence was found for an effect of lipid-regulating agents on probable (`r probad_text`) and possible (`r possad_text`) Alzheimer's disease when compared to no treatment, with the sole exception of fibrates on probable Alzheimer's disease (`r probad_fib_text`).

@@

In almost all cases, the use of imputed data resulted in a marginal attenuation of the effects observed when using a complete case analysis. It should be noted that due to the large amount of missing data (e.g., `r ldltext` were missing a baseline LDL cholesterol measure), the number of participants included in the complete case analysis was substantially smaller than that included when using imputed data. In this case, though the overall position of the effect estimates does not change substantially when using the imputed dataset, there is a noticeable gain in power.[@sterne2009]

@@

Lipid-regulating agents showed little evidence of an association with probable and possible Alzheimer's disease when compared to no treatment but were associated with increased risk of the all-cause dementia, vascular dementia and other dementias diagnoses. The estimate observed in each case was driven by the effects observed in the statin subgroup because a substantial majority of participants prescribed an LRA were prescribed a statin. For the other drug classes, no association was found with any outcome, with two exceptions. Ezetimibe was associated with increased risk of vascular and other dementias, while fibrates were associated with increased risk of all-cause dementia and probable Alzheimer's disease.

@@

<!----------------------------------------------------------------------------->
(ref:smeethComparison-cap) __Comparison of code lists__ - Sankey diagram comparing the codes used in this analysis with those used in the Smeeth _et al_ paper.[@smeeth2009] The outcomes and number of codes contributing to each are presented (the Smeeth _et al_ outcomes are on the right-hand side of the figure). The joining lines showing the overlap between the categories in the two analyses.

(ref:smeethComparison-scap) Comparison of code lists

@@

This analysis serves to illustrative the importance of the code lists chosen to define the outcomes of interest in EHR, particularly if they are used to define competing outcomes (e.g., AD vs non-AD dementia). The different codes used by Smeeth _et al._, in addition to an analytical approach that adjusted for covariates defined after the index date, may go some way to explaining why my analysis obtained different results despite the substantial overlap in the data sources used.

@@

* I found little evidence for an effect of lipid-regulating agents on probable or possible Alzheimer’s disease. However, lipid-regulating agent use was associated with an increased risk of all-cause, vascular and other dementias. In all cases, the estimated associations for the "any LRA" analyses were driven by those observed in the large statin subgroup.

### 2021-12-16 06-Chp-IPD-Analysis.Rmd
End of day:

As such, the aims of this analysis are two fold. Firstly, I plan to perform an IPD meta-analysis across identified cohorts to examine the impact of participant age-at-measurement and sex on the relationship between lipids and dementia risk. Secondly, I aim to expand the evidence base for the effect of lipids on dementia outcomes by obtaining estimates from previously unanalysed cohorts available via the Dementia Platform UK, a large consortium of dementia cohorts.

@@

Eligible data sources for this analysis were prospective cohort studies (see Section \@ref(ipd-apply-access) for details on how these studies were identified). Data sources which were cross-sectional, either by design or due to the available data (e.g. a study recorded data on participants at multiple time-points or "waves", but only data from a single wave could be accessed) were excluded. Similarly, due to the time and cost restraints within the scope of my thesis project, studies making use of population-level electronic health records were ineligible. These studies are problematic in that they often require extensive project proposals in order to gain access to the data.

@@

I considered four blood lipid fractions as part of this analysis, namely: total cholesterol (TC), low-density lipoprotein cholesterol (LDL-c), high-density lipoprotein cholesterol (HDL-c) and triglycerides (TG). Cohorts were eligible for inclusion if they contained data on at least one of these lipid fractions,  recorded as a continuous variable (i.e. studies with binary "hypercholesteromemia" would be excluded).

@@

The second approach focused on incorporating relevant, previously-unanalysed data into the analysis, thus providing additional evidence on the relationship between blood lipids and dementia risk. This was achieved through the Dementia Platform UK (DPUK), a collaborative grouping of existing dementia cohorts established by the Medical Research Council which works with data owners to make their data readily accessible for secondary analysis.[@bauermeister2020] It provides access to 42 cohorts with over 3 million participants, and makes use of a central streamlined application process for all cohorts, with the intent of making it easier to access data from existing data sources.

@@

(ref:cohortFlowchart-cap) __Flowchart detailing sources of data for the IPD analysis__ The number of cohorts approached and the reasons for exclusion are stratified by identification method (systematic review vs DPUK).

@@

<!-- TODO Be clear on the different types of databases, e.g. those where users are followed up regularly vs. those where participants are captured via EHR-->

@@

The Caerphilly Prospective Study (CaPS) is a longitudinal study of men in South Wales, UK.[@zotero-15398;@elwood2013] Blood lipids (TC, LDL-c, HDL-c and TG) were measured at baseline in 1979-1983, and from Phase III (1989-1993) onwards, a battery of cognitive tests were introduced. Dementia outcomes were determined using data obtained during Phase V (2002-2004), giving between 19-25 years of follow-up. The cohort contains data on dementia outcomes sub-classified as vascular and non-vascular dementia. Data was available for all covariates of interest except for ethnicity, BMI, prevalent IHD and _Apo_$\mathcal{E}4$ status. While the CaPS study collected data on height/weight (from which BMI can be calculated as $\frac{weight}{height^2}$) and prevalent heart disease, these were not available from the DPUK version of the CaPS data.

@@

The Whitehall II study is a prospective cohort study of men and women recruited between 1985 and 1989 from the civil service in London.[@marmot2005] Lipids measurements were available from the third wave of the study, conducted between 1991-1993. The cohort is linked with the Hospital Episode Statistics (HES) database, a database containing details of participant events at NHS hospitals in England. which was used to capture dementia outcomes.[@zotero-15403] HES data was available up to March 2015, providing between 22-24 years of follow-up. Information was available on dementia type, which was classified as all-cause dementia, Alzheimer's disease or vascular dementia. This data source contained details on all covariates of interest except for _Apo_$\mathcal{E}4$ status.

@@

Based on available data across cohorts, smoking was classified as never/ever/current, while alcohol consumption was classified as never/ever. Ever use in this case refers to participants who do not currently smoke/drink alcohol, but did so in the past. Education was categorised into 4 levels (None, O-levels, A-levels, Degree). BMI was treated as a continuous variable while presence of vascular co-morbidities was treated as dichotomous.

A key consideration in the definition of covariates across cohorts was the classification of age. The Whitehall II study, the largest cohort to which I had access, only shared age data in five-year age bands (e.g. 40-44, 45-49, etc.). To ensure comparability across the cohorts, I created identical categories in the CaPS and EPIC data. This grouped age variable was then used in all subsequent analyses.

@@

For the consideration of vascular dementia, I was only able to explore the effect of age, as only the CaPS and Whitehall cohorts contained details on vascular dementia as an outcome. As discussed above, the CaPS data contained a single sex excluding it from a exposure-sex analysis, leaving Whitehall II as the sole eligible study.

@@

This analysis requested data from `r n_applied$total` data sources, but only obtained data from three, all of which were based in the United Kingdom. No evidence for an effect of lipids on the risk of dementia or related outcomes was identified, except for a harmful association of raised triglycerides with risk of vascular dementia. Similarly, there was weak evidence for an interaction of the effect of lipid levels on dementia outcomes with participants' age (grouped into 5 year bands) or sex.

@@

The obstacles to data access described above are in theory what the DPUK was built to address. However, even with the help of the streamlined application process afforded by the DPUK, accessing sufficient data was a challenge. The response rate among DPUK cohorts a year after application was just 50%. In addition, some cohorts responded saying that the proposed study question was already under investigation by another group, and that they would not share the data on this basis. In light of this, a centralised database of ongoing analyses being performed using DPUK data would be of enormous help, reducing research waste and providing opportunities for collaboration. Finally, the DPUK process would be aided by a clearer distinction between those cohorts that are "DPUK native" (i.e. where a copy of the data is already held on DPUK servers) versus externally hosted, seeing as the response time for externally hosted cohorts is likely to be much longer.

@@

Finally, this analysis provides new evidence on a previously unexplored outcome (vascular dementia), adding to the extremely small evidence base for this outcome identified by the systematic review presented in Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading). This is relevant even in the previously analysed WhiteHall II cohort because the previous analysis by Tynkkynen _et al._ did not report on this outcome.

@@

In hindsight, attempting to undertake a large-scale IPD meta-analysis as part of a larger PhD project may have been overly ambitious. Data harmonization between cohorts in an IPD analysis is an often under-appreciated challenge,[@levis2021] and in line with this, data cleaning for this analysis took substantially longer than expected. While the cohort response rate was substantially lower that expected, given the time and resources required for the cleaning and harmonisation of data from just three cohorts, a situation in which all 37 cohorts responded positively would have been logistically challenging within the scope of a PhD.

@@

- I discussed potential reasons for the low response rate, and explored other limitations of this analysis. I highlighted the contribution of this work to the wider topic via the analysis of previously unexplored cohorts (CaPS & EPIC) and outcomes (vascular dementia). Finally, I recommended that future research could formally investigate the impact of the characteristics of the requesting researcher on data access rates.

>>>>

As such, the aims of this analysis are two-fold. Firstly, I plan to perform an IPD meta-analysis across identified cohorts to examine the impact of participant age-at-measurement and sex on the relationship between lipids and dementia risk. Secondly, I aim to expand the evidence base for the effect of lipids on dementia outcomes by obtaining estimates from previously unanalysed cohorts available via the Dementia Platform UK, a large consortium of dementia cohorts.

@@

Eligible data sources for this analysis were prospective cohort studies (see Section \@ref(ipd-apply-access) for details on how these studies were identified). Data sources which were cross-sectional, either by design or due to the available data (e.g., a study recorded data on participants at multiple time-points or "waves", but only data from a single wave could be accessed) were excluded. Similarly, due to the time and cost restraints within the scope of my thesis project, studies making use of population-level electronic health records were ineligible. These studies are problematic in that they often require extensive project proposals in order to gain access to the data.

@@

I considered four blood lipid fractions as part of this analysis, namely: total cholesterol (TC), low-density lipoprotein cholesterol (LDL-c), high-density lipoprotein cholesterol (HDL-c) and triglycerides (TG). Cohorts were eligible for inclusion if they contained data on at least one of these lipid fractions,  recorded as a continuous variable (i.e., studies with binary "hypercholesterolemia" exposure would be excluded).

@@

The second approach focused on incorporating relevant, previously unanalysed data into the analysis, thus providing additional evidence on the relationship between blood lipids and dementia risk. This was achieved through the Dementia Platform UK (DPUK), a collaborative grouping of existing dementia cohorts established by the Medical Research Council which works with data owners to make their data readily accessible for secondary analysis.[@bauermeister2020] It provides access to 42 cohorts with over 3 million participants, and makes use of a central streamlined application process for all cohorts, with the intent of making it easier to access data from existing data sources.

@@

(ref:cohortFlowchart-cap) __Flowchart detailing sources of data for the IPD analysis__ - The number of cohorts approached and the reasons for exclusion are stratified by identification method (systematic review vs DPUK).

@@

<!-- TODO Be clear on the different types of databases, e.g., those where users are followed up regularly vs. those where participants are captured via EHR-->

@@

The Caerphilly Prospective Study (CaPS) is a longitudinal study of men in South Wales, UK.[@zotero-15398;@elwood2013] Blood lipids (TC, LDL-c, HDL-c and TG) were measured at baseline in 1979-1983, and from Phase III (1989-1993) onwards, a battery of cognitive tests was introduced. Dementia outcomes were determined using data obtained during Phase V (2002-2004), giving between 19-25 years of follow-up. The cohort contains data on dementia outcomes sub-classified as vascular and non-vascular dementia. Data was available for all covariates of interest except for ethnicity, BMI, prevalent IHD and _Apo_$\mathcal{E}4$ status. While the CaPS study collected data on height/weight (from which BMI can be calculated as $\frac{weight}{height^2}$) and prevalent heart disease, these were not available from the DPUK version of the CaPS data.

@@

The Whitehall II study is a prospective cohort study of men and women recruited between 1985 and 1989 from the civil service in London.[@marmot2005] Lipid measurements were available from the third wave of the study, conducted between 1991-1993. The cohort is linked with the Hospital Episode Statistics (HES) database, a database containing details of participant events at NHS hospitals in England. which was used to capture dementia outcomes.[@zotero-15403] HES data was available up to March 2015, providing between 22-24 years of follow-up. Information was available on dementia type, which was classified as all-cause dementia, Alzheimer's disease or vascular dementia. This data source contained details on all covariates of interest except for _Apo_$\mathcal{E}4$ status.

@@

Based on available data across cohorts, smoking was classified as never/ever/current, while alcohol consumption was classified as never/ever. Ever use in this case refers to participants who do not currently smoke/drink alcohol but did so in the past. Education was categorised into 4 levels (None, O-levels, A-levels, Degree). BMI was treated as a continuous variable while presence of vascular co-morbidities was treated as dichotomous.

A key consideration in the definition of covariates across cohorts was the classification of age. The Whitehall II study, the largest cohort to which I had access, only shared age data in five-year age bands (e.g., 40-44, 45-49, etc.). To ensure comparability across the cohorts, I created identical categories in the CaPS and EPIC data. This grouped age variable was then used in all subsequent analyses.

@@

For the consideration of vascular dementia, I was only able to explore the effect of age, as only the CaPS and Whitehall cohorts contained details on vascular dementia as an outcome. As discussed above, the CaPS data contained a single sex which meant it was excluded from the exposure-sex analysis, leaving Whitehall II as the sole eligible study.

@@

This analysis requested data from `r n_applied$total` data sources, but only obtained data from three, all of which were based in the United Kingdom. No evidence for an effect of lipids on the risk of dementia or related outcomes was identified, except for a harmful association of raised triglycerides with risk of vascular dementia. Similarly, there was weak evidence for an interaction of the effect of lipid levels on dementia outcomes with participants' age (grouped into 5-year bands) or sex.

@@

The obstacles to data access described above are in theory what the DPUK was built to address. However, even with the help of the streamlined application process afforded by the DPUK, accessing sufficient data was a challenge. The response rate among DPUK cohorts a year after application was just 50%. In addition, some cohorts responded saying that the proposed study question was already under investigation by another group, and that they would not share the data on this basis. In light of this, a centralised database of ongoing analyses being performed using DPUK data would be of enormous help, reducing research waste and providing opportunities for collaboration. Finally, the DPUK process would be aided by a clearer distinction between those cohorts that are "DPUK native" (i.e., where a copy of the data is already held on DPUK servers) versus externally hosted, seeing as the response time for externally hosted cohorts is likely to be much longer.

@@

Finally, this analysis provides new evidence on a previously unexplored outcome (vascular dementia), adding to the extremely small evidence base for this outcome identified by the systematic review presented in Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading). This is relevant even in the previously analysed Whitehall II cohort because the previous analysis by Tynkkynen _et al._ did not report on this outcome.

@@

In hindsight, attempting to undertake a large-scale IPD meta-analysis as part of a larger PhD project may have been overly ambitious. Data harmonization between cohorts in an IPD analysis is an often under-appreciated challenge,[@levis2021] and in line with this, data cleaning for this analysis took substantially longer than expected. While the cohort response rate was substantially lower than expected, given the time and resources required for the cleaning and harmonisation of data from just three cohorts, a situation in which all 37 cohorts responded positively would have been logistically challenging within the scope of a PhD.

@@

- I discussed potential reasons for the low response rate and explored other limitations of this analysis. I highlighted the contribution of this work to the wider topic via the analysis of previously unexplored cohorts (CaPS & EPIC) and outcomes (vascular dementia). Finally, I recommended that future research could formally investigate the impact of the characteristics of the requesting researcher on data access rates.
